Improving Item-Item Similarity Estimation during Collaborative Filtering

### 2014-03-25 paper/paper.tex
Finished linear techniques section. Added graph. Going to bed

\begin{align}
ModSim(Sim(A, B), n) \approx lim_{n\to\infty}Sim(A, B)
\end{align}

@@

In our first approach, we model true and observed similarity probabilistically. The
goal is to use the observed similarity score directly to approximate the true similarity.
In this way, we can consider an abstact $TrueSim$ which ignores the details of 
$TrueSimRand$ or $TrueSimPref$. Similarly, we consider $Sim_{n}$ and $GoldSim$ to be
the generic observed similarity scores. Let $Y$ be a random variable representing 
the $TrueSim$ of a randomly chosen pair of items. Let $X_{n}$ be a random variable 
representing the $Sim_{n}$ of a pair of items when $n$ common users are observed.

@@

The model suggests that there is a linear correlation between $X_n$ ($Sim_n$)
and $Y$ ($TrueSim$). Moreover, the best predicted $TrueSim$ is a linear 
combination of the mean and the observed similiarity. Since $\sigma_{err,n}^2$ is 
nonnegative, the slope is always less than or equal to one, meaningthat the 
observed distance from $\mu$ (the average true similarity) overestimates the 
true distance. This is reasonable, since there is a prior that expects the true 
similarity to be around $\mu$.

This model requires calculating $\sigma_{err,n}^2$ for every value of $n$ 
separately. To gain a more general model, which relates the $\sigma_{err, n}$'s, 
we model $\sigma_{err,n}^2$ as a function of $n$. Recall that $\sigma_{err,n}^2$ 
is the noisiness of the observed $Sim_n$ about $TrueSim$. Since the $TrueSim$ is
the $Sim$ when all users that will rate the items have rated them, $Sim_n$ can be
thought of as sampling $n$ users from the set used to calculate $TrueSim$. 
Although $Sim$ is likely more complicated than an average, the Central Limit 
Theorem motivates the intuition that the variance of $Sim$ would decrease as the 
$\sqrt{n}$. Thus we model $\sigma_{err,n}^2$ as:
\begin{align}
\sigma_{err,n}^2 = \frac{\alpha}{\sqrt{n}},
\end{align}
yielding the single parameter model:
\begin{align}
\left(Y - \mu\right) = \frac{\sigma_{s}^2}{\sigma_{s}^2+\frac{\alpha}{\sqrt{n}}}
\left(X_n-\mu\right)
\end{align}

\subsection*{Technique}
The parameters $\mu$, $\sigma_{s}^2$, and $\sigma_{err,n}^2$ can be approximated 
using the method of moments over training data. $\mu$ and $\sigma_{s}$ are
calculated

Based on this maodel, we try three different techniques to predict the true similarity from a
measured similarity. The techniques vary in their fidelity to the model. For
each, we limit our data to all pairs of items that have enough common users that
we can approximate $TrueSim \approx Sim$. Then we pick subsets of the common
users of various sizes $n$ and calculate $Sim_{n}$. With examples of $Sim$, $n$
and the resulting $TrueSim$, we can implement supervised learning.

The first technique is the least faithful to the model. It simply runs a number
of linear regressions between $Sim$ and $TrueSim$ for each $n$. This ignores the 
data-specific parameters $\mu$, $\sigma_{s}$ and $\sigma_{err,n}$ to find the best 
linear fit. It also treats each $n$ completely independently. This technique 
necessarilty has the lowest training error of all linear techniques, potentially 
at the risk of overfitting.

The second technique is more general. It approximates $\mu$ and $\sigma_{s}$
using the method of moments on the training data of all $TrueSim$ scores. Then,
$\sigma_{err,n}$ is also approximated using a method of moments on the training
data of all $Sim_n$ for each $n$ separately. Here $\sigma_{err,n}$'s are also
considered independently of each other.

The third and most general technique models $\sigma_{err,n}$ as a function of
$n$ using equation 19. As in the previous technique, $\mu$, $\sigma_{s}$ and each
$\sigma_{err,n}$ are approximated using the method of moments over the training
data. Then $\alpha$ is calculated that minimizes the sum of squared errors
between $\sigma_{err,n}^2$ and $\frac{\alpha}{\sqrt{n}}$. This $\alpha$ is used to
construct the three-parameter prediction function of equation 19.

>>>>

In the following sections, we present two approaches to better approximate the
true similarity given a limited number of common reviewers. Both approaches are
motivated by probabilistic models with distributions grounded in the data. The
first approach leverages the observed similarity score and the number of common
users to calculate the true similarity. The second approach leverages the
rating-vectors themselves.

@@

In our first approach, we model the probability distributions of true and 
observed similarity. The goal is to use the observed similarity score directly 
to approximate the true similarity. In this way, we can consider an abstact 
$TrueSim$ which ignores the details of $TrueSimRand$ or $TrueSimPref$. 
Similarly, we consider $Sim_n$ to be a generic observed similarity score when
the two itmes have $n$ common reviewers. Now, let $Y$ be a random variable 
representing the $TrueSim$ of a randomly chosen pair of items. Let $X_{n}$ be a
random variable representing the $Sim_{n}$ of a pair of items when $n$ common 
users are observed.

@@

The model suggests that there is a linear correlation between $X_n$ ($Sim_n$)
and $Y$ ($TrueSim$). Moreover, the best predicted $TrueSim$ is a linear 
combination of the average true similarity and the observed similiarity. Since $\sigma_{err,n}^2$ is 
nonnegative, the slope is always less than or equal to one, meaning that the 
observed distance from $\mu$ (the average true similarity) overestimates the 
true distance. This is reasonable, since prior suggests that the true 
similarity is likely to be near $\mu$.\footnote{It is worth noting that despite
the resemblance to Slope One Predictors \cite{Lemire1998}, this is a fundamentally 
different technique. Here, we have a linear relationship between observed 
scores and true scores, whereas the Slope One model seeks a linear relationship 
between the ratings a user gives to one items versus another.}

This model requires a different parameter $\sigma_{err,n}^2$ for each $n$.
To gain a more generality, we model $\sigma_{err,n}^2$ as a function of $n$
instead of considering them independent of each other. Recall that 
$\sigma_{err,n}^2$ is the noisiness of the observed $Sim_n$ about $TrueSim$. 
Since the $TrueSim$ is the $Sim$ when all users that will rate the items 
have rated them, $Sim_n$ can be thought of as sampling $n$ users from the 
set used to calculate $TrueSim$. Although $Sim$ is likely more complicated 
than an average, the Central Limit Theorem motivates the intuition that the 
variance of $Sim$ would decrease as the sqaure root of $n$. Thus we model 
$\sigma_{err,n}^2$ as:
\begin{align}
\sigma_{err,n}^2 = \frac{\alpha}{\sqrt{n}},
\end{align}
yielding the single parameter model:
\begin{align}
\left(Y - \mu\right) = \frac{\sigma_{s}^2}{\sigma_{s}^2+\frac{\alpha}{\sqrt{n}}}
\left(X_n-\mu\right)
\end{align}

\subsection*{Technique}
While it may be impossible to get $TrueSim$ for any pair, we
can get close by calculating $Sim_n$ when a large number of users have rated
both items. Formally, we define $GoldSim$ = $Sim_N$ where $N$ is a number 
large enough that $GoldSim \approx TrueSim$.

With a measurable $GoldSim$ as our goal, we can apply supervised learning 
techniques. First, we find pairs of items that have at least $N$ common users 
and calculate their $GoldSim$. Next we withold some of the users to produce 
estimated $Sim_n$ for $n=1,...,N$. By doing this over many pairs of items, we 
produce pairs of ($Sim_n$, $TrueSim$) that can be used for training and testing.

Now, the parameters $\mu$, $\sigma_{s}^2$, and $\sigma_{err,n}^2$ can be 
approximated using the method of moments over the training data. The values of 
$\sigma_{err,n}^2$ are fit to the function $\frac{\alpha}{\sqrt{n}}$ and the $\alpha$ 
that minimizes the sum of squared error is computed.

The following graph compares the results of three linear fits, with varying
degrees of fidelity to the probabilistic model. The most faithful fit uses
$\mu$, $\sigma_{s}^2$ and $\alpha$ as calculated above to predict the $GoldSim$
using Eqn (18). The next fit uses the calculated values of $\sigma_{err,n}^2$ directly,
following Eqn (16) for predictions. The third fit simply uses a series of linear
regressions, one for each $n$. This model only assumes that for each $n$,
$Sim_n$ and $GoldSim$ are linearly correlated.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.7\textwidth]{LinearTechniques.png}
	\caption{Comparison of different linear models. For each number of reviewers
    in common $n$, the graphs represent the average squared error of predicted
    $GoldSim$ calculated from the observed $Sim_n$. As a baseline, the technique
    of guessing the average $GoldSim$ is included.}
    \label{fig:Linear}
\end{figure}

The linear regression technique trades the generality of the model for a better
fit on the training data. As the results show, the trade-off is worth it.
